---
title: "codeProject"
output: html_document
date: "2025-06-13"
---

load required libraries for data manipulation, visualization, and reporting
```{r}
library(tidyverse)
library(broom)      
library(knitr)      
library(scales)
library(dplyr)
```

```{r}
# load accident and vehicle datasets
cars_accidents_data = read.csv("data/CarsAccidentsData.csv")
cars_data = read.csv("data/CarsDatanew.csv")

# define alert and active safety system column names
alert_systems_cols = c("BLIND_SPOT_DETECTION","FRNT_COLLISION_WARNING", "LANE_DEPARTURE_WARNING")
active_systems_cols = c("CRASH_IMMINENT_BRAKE", "DYNAMIC_BRAKE_SUPPORT", "NHTSA_ESC")


```


Preprocess the vehicle dataset to enable normalization for modeling
```{r}

# filter and clean the vehicle dataset for modeling
all_systems = c(alert_systems_cols, active_systems_cols)


# convert system values to numeric encoding:
convert_system_values = function(system_value) {
  case_when(
    is.na(system_value) ~ 0,
    tolower(system_value) %in% c("standard", "standart", "std", "available") ~ 1,
    tolower(system_value) %in% c("optional", "option") ~ 2,
    tolower(system_value) %in% c("none", "n/a", "not available", "") ~ 0,
    TRUE ~ 0
  )
}

set.seed(123)

cars_clean_data  = cars_data %>%
  filter(MODEL_YR >= 2011 & MODEL_YR <= 2023) %>%   # keep only recent model years
  mutate_at(all_systems, convert_system_values) %>%         
  mutate_at(all_systems, ~ ifelse(. == 2, sample(c(0, 1), size = length(.), replace = TRUE), .)) %>%  
  distinct(MAKE, MODEL, MODEL_YR, .keep_all = TRUE) # remove duplicates by vehicle identity



```


calculate and display the distribution of vehicles by number of installed safety systems
```{r}

systems_by_count = cars_clean_data %>%
  mutate(system_count = rowSums(.[, all_systems] == 1)) %>%  # count how many systems are installed (value == 1)
  group_by(system_count) %>%
  summarise(
    Count = n(),
    Percent = round(100 * Count / nrow(.), 2) / 100,
    .groups = "drop"
  ) %>%
  arrange(system_count)

systems_by_count
```


 generate a table showing the percentage of each unique safety system combination
```{r}

# extract only the relevant safety system columns and convert system values to binary
binary_system_matrix = cars_clean_data[, all_systems]
binary_system_matrix = as.data.frame(lapply(binary_system_matrix, function(system_value) as.integer(system_value == 1)))

# create a binary string for each vehicle
binary_system_matrix$system_combo_string = apply(binary_system_matrix, 1, paste0, collapse = "")

# summarize counts and percentages of each unique combination
system_combo_summary  = binary_system_matrix %>%
  group_by(system_combo_string) %>%
  summarise(
    Count = n(),
    Percent = round(100 * Count / nrow(binary_system_matrix), 2),
    .groups = "drop"
  ) %>%
  arrange(desc(Count))

# split combo string into individual system columns
combo_binary_split = do.call(rbind, strsplit(system_combo_summary$system_combo_string, split = ""))
combo_binary_split = as.data.frame(combo_binary_split)
colnames(combo_binary_split) = all_systems

# combine split matrix with summary table
car_safey_combinations = cbind(combo_binary_split, system_combo_summary[, c("Count", "Percent")])
car_safey_combinations = car_safey_combinations %>%
  mutate(across(all_of(all_systems), as.integer))

car_safey_combinations
```


percentage of vehicles that include: alert systems, active systems, or both
```{r}

GroupSummary = cars_clean_data %>%
  mutate(
    has_alert = rowSums(.[, alert_systems_cols] == 1) > 0,
    has_active = rowSums(.[, active_systems_cols] == 1) > 0,
    group = case_when(
      has_alert & has_active ~ "Both",
      has_alert & !has_active ~ "Alerts Only",
      !has_alert & has_active ~ "Active Only",
      TRUE ~ "None"
    )
  ) %>%
  group_by(group) %>%
  summarise(
    Count = n(),
    Percent = round(100 * Count / nrow(cars_clean_data), 2),
    .groups = "drop"
  ) %>%
  arrange(desc(Percent))

GroupSummary
```


 analyze distribution of vehicles based on alert system configurations
```{r}

# extract alert system columns and convert to binary
alert_system_matrix  = cars_clean_data[, alert_systems_cols]
alert_system_matrix [] = lapply(alert_system_matrix , function(system_value) as.integer(system_value == 1))

# create a binary string representing each alert system combination
alert_system_matrix$alert_combo_string = apply(alert_system_matrix , 1, paste0, collapse = "")

# summarize frequency and percentage of each unique alert system combination
alert_combo_summary = alert_system_matrix  %>%
  group_by(alert_combo_string) %>%
  summarise(
    Count = n(),
    alert_system_precent = round(100 * Count / nrow(alert_system_matrix ), 2) / 100,
    .groups = "drop"
  ) %>%
  arrange(desc(Count))

# create final alert system group table
alert_system_group = alert_combo_summary %>%
  mutate(
    BLIND_SPOT_DETECTION = as.integer(substr(alert_combo_string, 1, 1)),
    FRNT_COLLISION_WARNING = as.integer(substr(alert_combo_string, 2, 2)),
    LANE_DEPARTURE_WARNING = as.integer(substr(alert_combo_string, 3, 3)),
    combo_alert_systems = paste0(BLIND_SPOT_DETECTION, FRNT_COLLISION_WARNING, LANE_DEPARTURE_WARNING)
  )

# reorder columns in final alert system summary table
alert_system_group = alert_system_group[, c("BLIND_SPOT_DETECTION", "FRNT_COLLISION_WARNING", "LANE_DEPARTURE_WARNING",  "combo_alert_systems", "Count", "alert_system_precent")]

```

end of all normalization and preprocessing steps
----------------------------------------------------------------
start of ploting the data

```{r}

set.seed(123) 

# randomly assign values (0 or 1) for alert systems marked as "optional" (value == 2)
cars_accidents_data = cars_accidents_data %>%
  mutate(across(all_of(alert_systems_cols), ~ ifelse(.x == 2, sample(c(0, 1), size = length(.x), replace = TRUE), .x)))

# same for active systems
cars_accidents_data = cars_accidents_data %>%
  mutate(across(all_of(active_systems_cols), ~ ifelse(.x == 2, sample(c(0, 1), size = length(.x), replace = TRUE), .x)))

# remove unnecessary or irrelevant columns
cars_accidents_data = cars_accidents_data %>% 
  dplyr::select(-any_of(c("AUTO_CRASH_NOTIFICATION", "ARS")))

# count how many alert systems are active per vehicle
cars_accidents_data$notification_count = rowSums(cars_accidents_data[, alert_systems_cols] == 1)


# Filter the accidents dataset to include only valid and relevant driving conditions
cars_accidents_data = cars_accidents_data %>%
  filter(
    DR_DRINK == 0, # driver was not under the influence of alcohol
    VSURCONDNAME == "Dry", # road surface condition is dry
    !grepl("(?i)Yes", SPEEDRELNAME), # exclude cases where speed was related to the crash
    !grepl("(?i)Unknown", SPEEDRELNAME), # exclude cases with unknown speed relation
    L_STATUSNAME == "Valid", # license status is valid
    IS_DRIVER_PRESENT == 1 # a driver was present in the vehicle
  )
```



```{r}


# summarize accidents by safety_count
accidents_by_safety_count = cars_accidents_data %>%
  group_by(SAFETY_COUNT) %>%
  summarise(num_accidents = n(), .groups = "drop") %>%
    filter(SAFETY_COUNT %in% 1:5) %>%  
  mutate(
    group_size = systems_by_count$Count[as.character(SAFETY_COUNT)],
    group_prob = systems_by_count$Percent[as.character(SAFETY_COUNT)],
    accidents_per_vehicle = num_accidents / group_size,
    accidents_per_percent = num_accidents / group_prob
  )

# plot actual number of accidents
ggplot(accidents_by_safety_count, aes(x = factor(SAFETY_COUNT), y = num_accidents)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Number of Accidents by Number of Safety Systems",
    x = "Number of Safety Systems",
    y = "Number of Accidents"
  ) +
  theme_minimal() +
  theme(
    plot.background = element_rect(fill = "white", color = NA)
  )
```

```{r}

# categorize vehicles based on presence of alert/active systems, then filter to only include severe crashes with at least one system

cars_accidents_data_combo_Group = cars_accidents_data %>%
  mutate(
    has_alert = rowSums(across(all_of(alert_systems_cols)) == 1) > 0,
    has_active = rowSums(across(all_of(active_systems_cols)) == 1) > 0,
    group = case_when(
      has_alert & has_active ~ "Both",
      has_alert & !has_active ~ "Alerts Only",
      !has_alert & has_active ~ "Active Only",
      TRUE ~ "None"
    ),
    is_severe = DEFORMEDNAME %in% c("Disabling Damage", "Functional Damage")
  ) %>%
  filter(group != "None", is_severe)

# summarize number of severe crashes by system group
accident_summary = cars_accidents_data_combo_Group %>%
  group_by(group) %>%
  summarise(SevereCrashCount = n(), .groups = "drop")

# merge with population proportions and normalize crash rate
accident_summary = accident_summary %>%
  inner_join(GroupSummary, by = "group") %>%
  mutate(
    PopulationShare = Percent / 100, # convert percent to proportion
    NormalizedRate = SevereCrashCount / PopulationShare # normalize crash count by population share
  ) %>%
  arrange(desc(NormalizedRate))

# plot normalized severe crash rate by system group
ggplot(accident_summary, aes(x = fct_reorder(group, NormalizedRate), y = NormalizedRate)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Normalized Severe Crash Rate by System Group",
    x = "System Group",
    y = "Severe Crashes / Population Share"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "none",
        plot.background = element_rect(fill = "white", color = NA)
        )
```


```{r}
# Function: compute accident rates for each alert system with 0 or 1 active system
get_alert_curve_one_active = function(alert_col, accident_data, population_data) {
  results = list()
  
  # Only include: no active system, or exactly one at a time
  combos = list(NULL)  # "None"
  for (sys in active_systems_cols) {
    combos[[length(combos) + 1]] = c(sys)
  }

  for (combo in combos) {
    combo_name = if (length(combo) == 0) "None" else combo[1]
    
    # Filter population: this alert system + exactly this active system (no extras)
    filtered_pop = population_data %>%
      filter(!!sym(alert_col) == 1,
             rowSums(across(all_of(combo), ~ . == 1)) == length(combo),
             rowSums(across(all_of(setdiff(active_systems_cols, combo)), ~ . == 1)) == 0)

    filtered_acc = accident_data %>%
      filter(!!sym(alert_col) == 1,
             rowSums(across(all_of(combo), ~ . == 1)) == length(combo),
             rowSums(across(all_of(setdiff(active_systems_cols, combo)), ~ . == 1)) == 0)

    accident_rate = if (nrow(filtered_pop) > 0) {
      100 * nrow(filtered_acc) / nrow(filtered_pop)
    } else {NA}

    results[[length(results) + 1]] = data.frame(
      Alert = alert_col,
      ActiveAdded = combo_name,
      AccidentRate = accident_rate
    )
  }

  bind_rows(results)
}

# Create alert combination column
cars_accidents_data$combo_alert_systems = apply(cars_accidents_data[, alert_systems_cols], 1, paste0, collapse = "")

# Filter accident cases, exclude extreme alert combos
accidents_rates = cars_accidents_data %>%
  filter(DEFORMEDNAME %in% c("Disabling Damage", "Functional Damage")) %>%
  filter(!combo_alert_systems %in% c("000", "111"))

# Run only over one-active-at-a-time combinations
final_df = bind_rows(lapply(alert_systems_cols, get_alert_curve_one_active,
                             accident_data = accidents_rates,
                             population_data = cars_clean_data)) %>%
  filter(!is.na(AccidentRate))

# Extract baseline when no active system was added
baseline_df = final_df %>%
  filter(ActiveAdded == "None") %>%
  select(Alert, BaselineRate = AccidentRate)

# Compute relative change from baseline
final_df = final_df %>%
  left_join(baseline_df, by = "Alert") %>%
  mutate(
    RelativeChangeFromNone = AccidentRate - BaselineRate
  ) %>%
  filter(ActiveAdded != "None")  # keep only additions

# Rename active systems for clarity
final_df$ActiveAdded = recode(final_df$ActiveAdded,
  "CRASH_IMMINENT_BRAKE" = "Crash\nImminent\nBrake",
  "DYNAMIC_BRAKE_SUPPORT" = "Dynamic\nBrake\nSupport",
  "NHTSA_ESC" = "Electronic\nStability\nControl"
)

# Final unified plot
ggplot(final_df, aes(x = ActiveAdded, y = RelativeChangeFromNone, color = Alert)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(size = 5) +
  labs(
    title = "Effect of Adding One Active System to Each Alert System",
    subtitle = "Each point shows change in accident rate compared to alert-only baseline",
    x = "Active System Added",
    y = "Relative Change from Baseline (%)",
    color = "Alert System",
    shape = "Alert System"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(size = 15, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text.x = element_text(size = 11, lineheight = 0.9),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA)
  )
```
end of plotng the data

-----------------------------------------------------------

modeling


```{r}

# assign group labels and identify severe crashes
cars_accidents_data_grouped = cars_accidents_data %>%
  mutate(
    has_alert = rowSums(across(all_of(alert_systems_cols)) == 1) > 0,
    has_active = rowSums(across(all_of(active_systems_cols)) == 1) > 0,
    group = case_when(
      has_alert & has_active ~ "Both",
      has_alert & !has_active ~ "Alerts Only",
      !has_alert & has_active ~ "Active Only",
      TRUE ~ "None"
    ),
    severe = DEFORMEDNAME %in% c("Disabling Damage", "Functional Damage")
  ) %>%
  left_join(GroupSummary, by = "group") %>%  # join with group population percentages
  mutate(
    weight = 1 / (Percent / 100) # inverse population weight
  )

# convert group to factor with "None" as the reference level for regression
cars_accidents_data_grouped$group = relevel(factor(cars_accidents_data_grouped$group), ref = "None")


# first model: Both vs Alerts Only
data_both_alerts = cars_accidents_data_grouped %>%
  filter(group %in% c("Both", "Alerts Only")) %>%
  mutate(is_both = ifelse(group == "Both", 1, 0))

model_both_vs_alerts = glm(severe ~ is_both, data = data_both_alerts,
                            weights = weight, family = "binomial")

coef_both_vs_alerts = tidy(model_both_vs_alerts) %>%
  filter(term != "(Intercept)") %>%
  mutate(Predictor = "Both vs Alerts Only")

# second model: Both vs Active Only
data_both_active = cars_accidents_data_grouped %>%
  filter(group %in% c("Both", "Active Only")) %>%
  mutate(is_both = ifelse(group == "Both", 1, 0))

model_both_vs_active = glm(severe ~ is_both, data = data_both_active,
                            weights = weight, family = "binomial")

coef_both_vs_active = tidy(model_both_vs_active) %>%
  filter(term != "(Intercept)") %>%
  mutate(Predictor = "Both vs Active Only")

# combine both models coefficients for visualization
coefs_combined = bind_rows(coef_both_vs_alerts, coef_both_vs_active) %>%
  mutate(
    ymin = estimate - 1.96 * std.error,
    ymax = estimate + 1.96 * std.error
  )

# plot effect of having Both systems vs each individual system
ggplot(coefs_combined, aes(x = Predictor, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Effect of Both Systems vs Each Alone",
    x = "Group Comparison",
    y = "Log Odds Estimate"
  ) +
  theme_minimal() + 
  theme(plot.background = element_rect(fill = "white", color = NA)
        )
```


```{r}

# Logistic regression model with normalized weights
model = glm(severe ~ group, data = cars_accidents_data_grouped,
             weights = weight, family = "binomial")

# Extract coefficients for Alerts Only and Active Only
coef_df = tidy(model) %>%
  filter(term %in% c("groupAlerts Only", "groupActive Only")) %>%
  mutate(
    Predictor = case_when(
      term == "groupAlerts Only" ~ "Alerts Only vs None",
      term == "groupActive Only" ~ "Active Only vs None"
    ),
    ymin = estimate - 1.96 * std.error,
    ymax = estimate + 1.96 * std.error
  )

# Plot coefficient estimates with confidence intervals
ggplot(coef_df, aes(x = Predictor, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = "Effect of Warning and Active Systems vs None",
    x = "Group Comparison",
    y = "Log Odds Estimate"
  ) +
  theme_minimal()


```

```{r}

# filter keep only disabling crashes with specific collision types
cars_accidents_data = cars_accidents_data %>%
  filter(DEFORMEDNAME %in% c("Disabling Damage", "Functional Damage"),
         TYPE_OF_COLLIOSION %in% c("Angle", "Front-to-Front", "Front-to-Rear","Rear-to-Side","Sideswipe - Same Direction", "Sideswipe - Opposite Direction")
         ) %>%
  filter(!combo_alert_systems %in% c("000", "111"))


alert_system_group = alert_system_group %>%
  filter(!alert_system_group$combo_alert_systems %in% c("000", "111"))


# join percentage distribution from alert system group, and compute inverse weights
cars_accidents_data = cars_accidents_data %>%
  left_join(
    alert_system_group %>% dplyr::select(combo_alert_systems, alert_system_precent),
    by = "combo_alert_systems"
  ) %>%
  mutate(
    weight_alert_system = 1 / alert_system_precent
  )


# logistic regression models for each collision type with interaction terms and normalize weight for each system/combintions of systems
model_front = glm(grepl("Front", TYPE_OF_COLLIOSION) ~
                     FRNT_COLLISION_WARNING + LANE_DEPARTURE_WARNING + BLIND_SPOT_DETECTION +
                     FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING +
                     FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION +
                     LANE_DEPARTURE_WARNING:BLIND_SPOT_DETECTION,
                   data = cars_accidents_data, family = "binomial", weights = weight_alert_system)

model_side = glm(grepl("Side", TYPE_OF_COLLIOSION) ~
                    FRNT_COLLISION_WARNING + LANE_DEPARTURE_WARNING + BLIND_SPOT_DETECTION +
                    FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING +
                    FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION +
                    LANE_DEPARTURE_WARNING:BLIND_SPOT_DETECTION,
                  data = cars_accidents_data, family = "binomial", weights = weight_alert_system)

model_angle = glm(grepl("Angle", TYPE_OF_COLLIOSION) ~
                     FRNT_COLLISION_WARNING + LANE_DEPARTURE_WARNING + BLIND_SPOT_DETECTION +
                     FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING +
                     FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION +
                     LANE_DEPARTURE_WARNING:BLIND_SPOT_DETECTION,
                   data = cars_accidents_data, family = "binomial", weights = weight_alert_system)


```

```{r}

# extract coefficients and tag with crash type
extract_coefs = function(model, crash_type) {
  tidy(model) %>%
    mutate(CrashType = crash_type)
}

# combine coefficients from all three logistic regression models
coef_table = bind_rows(
  extract_coefs(model_front, "Front"),
  extract_coefs(model_side,  "Side"),
  extract_coefs(model_angle, "Angle")
) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term = recode(term,
      "FRNT_COLLISION_WARNING" = "Front Warning",
      "LANE_DEPARTURE_WARNING" = "Lane Warning",
      "BLIND_SPOT_DETECTION" = "Blind Spot",
      "FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING" = "Front + Lane",
      "FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION" = "Front + Blind",
      "LANE_DEPARTURE_WARNING:BLIND_SPOT_DETECTION" = "Lane + Blind"
    ),
    lower = estimate - 1.96 * std.error,
    upper = estimate + 1.96 * std.error
  )

# filter significant effects based on p-value
significant = coef_table %>%
  filter(p.value < 0.05)

kable(significant, digits = 3, caption = "Significant Effects (p < 0.05)")

```


```{r}

# create a full system combination string (alert + active systems) for each row
car_safey_combinations = car_safey_combinations %>%
  mutate(
    combo_all_systems = paste0(BLIND_SPOT_DETECTION, FRNT_COLLISION_WARNING, LANE_DEPARTURE_WARNING,
                               CRASH_IMMINENT_BRAKE, DYNAMIC_BRAKE_SUPPORT, NHTSA_ESC)
  )


# join full system combination proportions into the accidents dataset
cars_accidents_data = cars_accidents_data %>%
  left_join(
    car_safey_combinations %>%
      dplyr::select(
        BLIND_SPOT_DETECTION,
        FRNT_COLLISION_WARNING,
        LANE_DEPARTURE_WARNING,
        CRASH_IMMINENT_BRAKE,
        DYNAMIC_BRAKE_SUPPORT,
        NHTSA_ESC,
        combo_all_systems,
        Percent
      ),
    by = c(
      "BLIND_SPOT_DETECTION",
      "FRNT_COLLISION_WARNING",
      "LANE_DEPARTURE_WARNING",
      "CRASH_IMMINENT_BRAKE",
      "DYNAMIC_BRAKE_SUPPORT",
      "NHTSA_ESC"
    )
  )


# rename and compute inverse weight based on full system combination proportion
cars_accidents_data = cars_accidents_data %>%
  rename(
    precent_all_systems = Percent
  ) %>%
   mutate(
    weight_all_systems = 1 / precent_all_systems
  )

```



extend the previously significant models by adding logical active safety systems, to test whether their inclusion can reduce the likelihood of severe crashes.
```{r}
model_front_extend = glm(grepl("Front", TYPE_OF_COLLIOSION) ~
                     FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING:CRASH_IMMINENT_BRAKE + # 011100
                     FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING:DYNAMIC_BRAKE_SUPPORT + # 011010
                     FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION:CRASH_IMMINENT_BRAKE, # 110100
                   data = cars_accidents_data, family = "binomial", weights = weight_all_systems)

model_side_extend = glm(grepl("Side", TYPE_OF_COLLIOSION) ~
                    BLIND_SPOT_DETECTION:NHTSA_ESC + # 100001
                    FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING:NHTSA_ESC,# 011001
                  data = cars_accidents_data, family = "binomial", weights = weight_all_systems)

model_angle_extend = glm(grepl("Angle", TYPE_OF_COLLIOSION) ~
                     FRNT_COLLISION_WARNING:CRASH_IMMINENT_BRAKE + # 010100
                     FRNT_COLLISION_WARNING:DYNAMIC_BRAKE_SUPPORT + # 010010
                     FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING:CRASH_IMMINENT_BRAKE + # 011100
                     BLIND_SPOT_DETECTION:FRNT_COLLISION_WARNING:NHTSA_ESC, # 110001
                   data = cars_accidents_data, family = "binomial", weights = weight_all_systems)
```

 
```{r}

# combine coefficients from all three logistic regression models
coef_table = bind_rows(
  extract_coefs(model_front_extend, "Front"),
  extract_coefs(model_side_extend, "Side"),
  extract_coefs(model_angle_extend, "Angle")
) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term = recode(term,
         "FRNT_COLLISION_WARNING:CRASH_IMMINENT_BRAKE:LANE_DEPARTURE_WARNING" = "Front + Lane + Imminent_Brake",
         "FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING:DYNAMIC_BRAKE_SUPPORT" = "Front + Lane + Dynamic_brake",
         "FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION:NHTSA_ESC" = "Front + Blind + ESC",
      .default = term  # משאיר את כל היתר ללא שינוי
    ),
    lower = estimate - 1.96 * std.error,
    upper = estimate + 1.96 * std.error
  )

# filter significant effects based on p-value
significant = coef_table %>%
  filter(p.value < 0.05)

knitr::kable(significant, digits = 3, caption = "Significant Effects (p < 0.05)")
```



```{r}


# create a summary table showing odds ratio reduction after adding active safety systems
effects_df = data.frame(
  Comparison = c("Front + Lane -> +Dynamic_brake","Front + Blind -> +ESC"),
  From_OR = exp(c(model_front$coefficients["FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING"], model_angle$coefficients["FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION"])),
  To_OR = exp(c(model_front_extend$coefficients["FRNT_COLLISION_WARNING:LANE_DEPARTURE_WARNING:DYNAMIC_BRAKE_SUPPORT"], model_angle_extend$coefficients["FRNT_COLLISION_WARNING:BLIND_SPOT_DETECTION:NHTSA_ESC"])),
  CrashType = c("Front",  "Angle")
) %>%
  mutate(
    PercentChange = (From_OR - To_OR) / From_OR * 100
  )

# define color palette per crash type
colors = c("Front" = "indianred" ,  "Angle" = "steelblue")

# plot the percent reduction in odds 
ggplot(effects_df, aes(x = Comparison, y = PercentChange, fill = CrashType)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(
    aes(label = paste0(round(PercentChange, 1), "%")),
    vjust = -0.3,
    size = 5
  ) +
  scale_fill_manual(values = colors) +
  scale_y_continuous(
    labels = percent_format(scale = 1),
    limits = c(0, 50)
  ) +
  labs(
    title = "Percent Reduction in Crash Odds After Adding Safety Systems",
    y = "Reduction",
    x = "System Addition",
    fill = "Crash Type"
  ) +
  theme(
    axis.text.x = element_text(angle = 15, hjust = 1),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 8)
  )

```